Twisted 11-dimensional supergravity

### 2021-10-21 intro.tex
intro changes

In this paper, we provide perturbative descriptions of all twists of eleven-dimensional supergravity (before turning on any $\Omega$-background).
%By `twists' we mean ones which arise from setting a bosonic ghost to take a nonzero value of a particular supertranslation in the 11-dimensional supersymmetry algebra. 
Thanks to a straightforward classification, which we recall below, there are essentially two types of twists which deform the flat background $\RR^{11}$: 
\begin{itemize}
\item The minimal (or holomorphic) twist. 
The resulting twisted theory is defined on $\CC^5 \times \RR$ and is holomorphic along $\CC^5$ and topological along the `time' direction~$\RR$. 
The twist is $SU(5)$ invariant and involves a fixed choice of a Calabi--Yau form on $\CC^5$. 
\item The non-minimal twist. 
The resulting twisted theory is defined on $\CC^2 \times \RR^7$ and is holomorphic along $\CC^2$ and topological along $\RR^7$.
The twist is $SU(2)$ invariant and involves a fixed choice of a hyper K\"ahler structure on $\CC^2$.  
\end{itemize}

In~\cite{SWspinor}, the last two authors used the pure spinor formalism to describe the {\em free field limit} of the minimal twist of 11-dimensional supergravity on flat space, following work of Cederwall~\cite{Ced-towards,Ced-11d}. In fact, the result of \cite{SWspinor} shows that the pure spinor formalism is compatible with twisting, so that the twist of any multiplet can be recovered from the algebraic geometry of a neighborhood of the corresponding point in the scheme of square-zero elements.
(In separate work~\cite{EagerHahner}, the free-field limit of the non-minimal twist was derived from the component-field multiplet as used in the physics literature.)
%\brian{also reference fabian and richard}
The resulting theory is $\ZZ/2$ graded; the grading is inherited from the totalization of ghost number and intrinsic parity in the original, untwisted, eleven-dimensional theory.
In \S\ref{sec:dfn} we introduce a full interacting theory in th Batalin--Vilkovisky (BV) formalism which has, as its free limit, the twist computed in \cite{SWspinor}.\footnote{One subtlety is that unlike the usual BV formalism which involves a $\ZZ$-grading by ghost number, the 11-dimensional theory is only $\ZZ/2$ graded.
All odd fields can therefore be considered as ``ghosts'' in the twisted theory.}
Our main conjecture is that this interacting BV theory is equivalent to the minimal twist of 11-dimensional supergravity on flat space.

\begin{conj}
The $11$-dimensional partially holomorphic-topological theory on $\CC^5 \times \RR$ that we will define in \S \ref{sec:dfn} is equivalent to twisted supergravity on $\RR^{11}$ where the bosonic ghost takes value the holomorphic supercharge. 
\end{conj}

%\brian{Ingmar, maybe you can comment on the status of this conjecture.}
In this paper we will provide evidence for this conjecture on a variety of fronts. 
In \S\ref{sec:susy}, we will show that this theory has all of the residual supersymmetries present after performing the holomorphic twist.

In \S\ref{sec:dimred}, we compute dimensional reductions and show that our model is compatible with descriptions of twists of lower dimensional twists of supergravity. 
For instance, reducing along a circle in a complex plane agrees with the conjectural description of the $SU(4)$ twist of type IIA supergravity of \cite{CLsugra}.

In \S\ref{sec:ads} we propose a description of supergravity in twisted versions of both the ${\rm AdS}_7$ and ${\rm AdS}_4$ backgrounds. 
We show that the residual symmetries of supergravity in these backgrounds are present in our twisted background.

Finally, we describe the non-minimal twist of 11-dimensional supergravity as a background of our theory on $\CC^5 \times \RR$. 
This non-minimal twist is invariant for the group $G_2 \times SU(2)$. 
We find a match with a conjectural description of this $G_2 \times SU(2)$ invariant twist formulated by Costello in \cite{CostelloM5} and further developed in \cite{RY}. 
%Surprisingly, at the twisted level these symmetries

@@

Our 11-dimensional theory is defined more generally on the product of manifolds 
\[
X \times S
\]
where $X$ is a Calabi--Yau five-fold and $S$ is a smooth oriented real one-dimensional manifold. 
The theory is of a holomorphic gravitational flavor as it describes ``partial'' deformations of complex structures along $X$. 
We will explain what we mean by this momentarily.

In our theory there is a field which encodes this partial deformation of complex structure on the Calabi--Yau manifold $X$.
It is an even field 
\[
\mu \in \Omega^{0,1} (X , \T_X) \otimes C^\infty(S) 
\]
where $\T_X$ denotes the holomorphic tangent bundle on $X$.
Locally, $\mu$ can be decomposed as Beltrami-like differential
\[
\mu = \mu^i_j (z,\zbar, t) \d \zbar_i \frac{\partial}{\partial z_j} .
\]
Physically speaking, $\mu$ is a component of the metric which survives in the twisted theory, see \S \ref{s:components}.

@@

There is a background where the equation of motion involving the $\mu, \gamma^{1,0}, \gamma^{1,2}$ fields reads
\begin{align*}
\dbar \mu + \frac12 [\mu, \mu] + \Omega^{-1} \vee \left(\del \gamma^{1,0} \wedge \del \gamma^{1,2} \right) & = 0 \\
%(\dbar + \mu) \gamma^{1,0} = 
%(\dbar + \mu) \gamma^{1,2} & = 0 
%\\
%\div \mu & = 0 \\
%\d_S \mu = \d_S \gamma^{1,0} = \d_S \gamma^{1,2} = & = 0 .
\end{align*}
Additionally, there are the conditions that $\mu$ preserve the holomorphic volume form on $X$ and that all fields are locally constant along the topological direction $S$.  
%In the last line $\d_S$ is the de Rham differential on $L$ and implies that $\mu, \gamma^{1,0}$, and $\gamma^{1,2}$ are locally constant along $L$. 
Notice that were it not for the presence of the last term in the first equation then this would be the integrability equation for $\mu$ to determines a complex structure on $X$.
%The second two equations resemble the conditions that $\gamma^{1,0}, \gamma^{1,2}$ be holomorphic with the complex structure determined by $\mu$. 
%Finally, the fourth equation constrains $\mu$ to preserve the holomorphic volume form on $X$.

If we work in a background where one of $\gamma^{1,0}$ or $\gamma^{1,2}$ is zero then we see that $\mu$ is exactly a deformation of complex structure along $X$. 
In terms of the 11-dimensional geometry, this is a background which describes deformations of the natural transverse holomorphic foliation (THF structure) on $X \times S$ which further preserves the holomorphic volume form along the leaves.
We further unpack the equations of motion in more general backgrounds in \S \ref{s:components}, but leave a complete study for future work.
%The equations of motion state that $\gamma$ is constant along $\RR$ and holomorphic along $\CC^5$. 
%Further, $\gamma$ is only defined up to the additional of a total holomorphic derivative $\gamma \sim \gamma + \del \beta$ where $\beta$ is a holomorphic function on $\CC^5$ which is constant along $\RR$.

@@

The simplest field configuration in the twisted theory is the flat background which corresponds to considering to our 11-dimensional theory on $\CC^5 \times \RR$ where we equip $\CC^5$ with the flat holomorphic volume form. 
In this case we find a striking relationship to a certain infinite-dimensional simple super Lie algebra studied by Kac and collaborators \cite{KacBible,KacE510}.

\begin{thm}
The global symmetry algebra of the 11-dimensional theory on $\CC^5 \times \RR$ is equivalent to a central extension of the exceptional super Lie algebra $E(5,10)$. 
\end{thm}

In particular, correlation functions involving observables of the 11-dimensional theory will be constrained by the infinite-dimensional symmetry algebra $E(5,10)$. 
Given our main conjecture that the interacting 11-dimensional BV theory on $\CC^5 \times \RR$ is the twist of supergravity, we obtain the following.

@@

Through the dimensional reduction of our 11-dimensional theory we will find a match with Costello and Li's descriptions of twists of 10-dimensional supergravity in terms of BCOV theory.
In Table \ref{diagram}, we provide a summary of the comparisons with twists of supergravity in lower dimensions we perform. The squigly arrows denote further twists, and the solid arrows denote various kinds of reductions.

\begin{center}
\begin{tikzcd}[row sep=large]
                                                                         &  \parbox{3cm}{\centering minimal twist of 11d} \arrow[ld, "\text{topological }S^1" '] \arrow[d, "\text{slab }" '] \arrow[rdd, "\text{CY3}" ' pos=.75] \arrow[rr, rightsquigarrow] \arrow[rd, "\text{holomorphic }S^1"] &                                                             &  \parbox{3cm}{\centering non-minimal twist of 11d} \arrow[d, "\text{topological }S^1"] \\
 \parbox{3cm}{\centering minimal twist of type IIA} \arrow[r, "\text{orbifold}"] &  \parbox{3cm}{\centering minimal twist of type I} \arrow[r, rightsquigarrow, crossing over]                                                                                                                                                                                               &  \parbox{3cm}{\centering SU(4) twist of type IIA} \arrow[r, rightsquigarrow] &  \parbox{3cm}{\centering SU(2) twist of type IIA}                                                 \\
                                                                         &                                                                                                                                                                                                                                                          & \parbox{3cm}{\centering minimal twist of 5d $\cN=1$}                   &                                                                                                
\end{tikzcd}
\end{center}
\label{diagram}

In \S \ref{s:su4red} we will show that the reduction along $\{0\} \times \RR \times \{0\} \subset \CC^4 \times \CC \times \RR$ is equivalent to the $SU(4)$ twist of type IIA supergravity. 
The topological string approach does not lead to a description of the minimal, or holomorphic, twist of type IIA supergravity. 
In \S \ref{s:su5IIA}, we describe the reduction along the line $\{0\} \times \RR \subset \CC^5 \times \RR$ to obtain a conjectural description of the holomorphic $SU(5)$ twist of type IIA supergravity. 
In \S \ref{s:Ired}, we show that the reduction along the interval $\{0\}\times [0,1]\subset \CC^5\times [0,1]$ with certain boundary conditions on the endpoints of the interval is equivalent to the twist of type I supergravity. We further observe in \S \ref{s:orbifold} that the twist of type I arises as the fixed points of a natural $\ZZ/2$ action on the minimal twist of IIA. Next, in \S \ref{s:CY3} we study the reduction along a Calabi-Yau 3 fold to obtain a conjectural description of the minimal twist of 5d $\cN=1$ supergravity. Finally, we show in \S \ref{s:nonmin} how the non-minimal twist of 11d supergravity arises as a further twist of our holomorphic twist.

\subsection*{Future work}

\subsection*{Acknowledgements}

>>>>

In this paper, we provide perturbative descriptions of all twists of eleven-dimensional supergravity (before turning on any $\Omega$-background).
%By `twists' we mean ones which arise from setting a bosonic ghost to take a nonzero value of a particular supertranslation in the 11-dimensional supersymmetry algebra. 
Thanks to a straightforward classification, which we recall below, there are essentially two types of twists which deform the flat background $\RR^{11}$: 
\begin{itemize}[leftmargin=\parindent,itemsep=\parskip]
\item The minimal (or holomorphic) twist. 
The resulting twisted theory is defined on $\CC^5 \times \RR$ and is holomorphic along $\CC^5$ and topological along the `time' direction~$\RR$. 
The twist is $SU(5)$ invariant and involves a fixed choice of a Calabi--Yau form on $\CC^5$. 
\item The non-minimal twist. 
The resulting twisted theory is defined on $\CC^2 \times \RR^7$ and is holomorphic along $\CC^2$ and topological along $\RR^7$.
The twist is $SU(2)$ invariant and involves a fixed choice of a hyperk\"ahler structure on $\CC^2$.  
\end{itemize}

In~\cite{SWspinor}, the last two authors used the pure spinor formalism to describe the {\em free field limit} of the minimal twist of eleven-dimensional supergravity on flat space, building on work of Cederwall~\cite{Ced-towards,Ced-11d} that constructed the eleven-dimensional theory in the pure spinor formalism. In fact, the result of~\cite{SWspinor} shows that the pure spinor formalism is compatible with twisting, so that the twist of any multiplet can be recovered from the algebraic geometry of a neighborhood of the corresponding point in the scheme of square-zero elements.
(In separate work~\cite{EagerHahner}, the free-field limit of the non-minimal twist was derived from the component-field multiplet as used in the physics literature.)
%\brian{also reference fabian and richard}
The resulting theory is $\ZZ/2$ graded; the grading is inherited from the totalization of ghost number and intrinsic parity in the original, untwisted, eleven-dimensional theory.
In \S\ref{s:dfn} we introduce a full interacting theory in the Batalin--Vilkovisky (BV) formalism, the free limit of which is the twist computed in \cite{SWspinor}.\footnote{One subtlety is that, unlike the usual BV formalism which involves a $\ZZ$-grading by ghost number, the eleven-dimensional theory is only $\ZZ/2$ graded.
All odd fields can therefore in some sense be considered as ``ghosts'' in the twisted theory.}
Our main conjecture is that this interacting BV theory is equivalent to the minimal twist of eleven-dimensional supergravity on flat space.

\begin{conj}
The eleven-dimensional partially holomorphic-topological theory on $\CC^5 \times \RR$ that we will define in \S \ref{s:dfn} is equivalent to twisted supergravity on $\RR^{11}$ where the bosonic ghost takes value the holomorphic supercharge. 
\end{conj}

%\brian{Ingmar, maybe you can comment on the status of this conjecture.}
In this paper we will provide evidence for this conjecture on a variety of fronts:
\begin{itemize}[leftmargin=\parindent,itemsep=\parskip]
\item
In \S\ref{sec:susy}, we will show that this theory has all of the residual supersymmetries present after performing the holomorphic twist. 
\item
In \S\ref{sec:dimred}, we compute dimensional reductions and show that our model is compatible with descriptions of twists of lower dimensional twists of supergravity. 
For instance, reducing along a circle in a complex plane agrees with the conjectural description of the $SU(4)$ twist of type IIA supergravity in~\cite{CLsugra}.
\item
In \S\ref{sec:ads} we propose a description of supergravity in twisted versions of both the ${\rm AdS}_7$ and ${\rm AdS}_4$ backgrounds. 
We show that the residual symmetries of supergravity in these backgrounds are present in our twisted background.
\item
Finally, we describe the non-minimal twist of eleven-dimensional supergravity as a background of our theory on $\CC^5 \times \RR$. 
This non-minimal twist is invariant for the group $G_2 \times SU(2)$. 
We find a match with a conjectural description of this $G_2 \times SU(2)$ invariant twist formulated by Costello in \cite{CostelloM5} and further developed in~\cite{RY}. 
\end{itemize}
%Surprisingly, at the twisted level these symmetries

@@

Our eleven-dimensional theory is defined more generally on the product of manifolds 
\[
X \times S
\]
where $X$ is a Calabi--Yau five-fold and $S$ is a smooth oriented real one-dimensional manifold. 
The theory is of a holomorphic gravitational flavor as it describes ``partial'' deformations of complex structures along $X$. 
We will explain what we mean by this momentarily.

In our theory there is a field which encodes this partial deformation of complex structure on the Calabi--Yau manifold $X$.
It is an even field 
\[
\mu^{1,1} \in \Omega^{0,1} (X , \T_X) \otimes C^\infty(S) 
\]
where $\T_X$ denotes the holomorphic tangent bundle on $X$.
Locally, $\mu^{1,1}$ can be decomposed as Beltrami-like differential
\[
\mu^{1,1} = \mu^i_j (z,\zbar, t) \d \zbar_i \frac{\partial}{\partial z_j} .
\]
Physically speaking, $\mu^{1,1}$ is a component of the metric which survives in the twisted theory; see~\S\ref{s:components}.

@@

There is a background where the equation of motion involving the fields $\mu^{1,1}$, $\gamma^{1,0}$,  and~$\gamma^{1,2}$ reads
\begin{align*}
\dbar \mu^{1,1} + \frac12 [\mu^{1,1}, \mu^{1,1}] + \Omega^{-1} \vee \left(\del \gamma^{1,0} \wedge \del \gamma^{1,2} \right) & = 0 \\
%(\dbar + \mu) \gamma^{1,0} = 
%(\dbar + \mu) \gamma^{1,2} & = 0 
%\\
%\div \mu & = 0 \\
%\d_S \mu = \d_S \gamma^{1,0} = \d_S \gamma^{1,2} = & = 0 .
\end{align*}
Additionally, there are the conditions that $\mu^{1,1}$ preserve the holomorphic volume form on $X$ and that all fields are locally constant along the topological direction $S$.  
%In the last line $\d_S$ is the de Rham differential on $L$ and implies that $\mu, \gamma^{1,0}$, and $\gamma^{1,2}$ are locally constant along $L$. 
Notice that this would be precisely the integrability equation for $\mu^{1,1}$ to determine a complex structure on $X$, were it not for the presence of the last term in the first equation.
%The second two equations resemble the conditions that $\gamma^{1,0}, \gamma^{1,2}$ be holomorphic with the complex structure determined by $\mu$. 
%Finally, the fourth equation constrains $\mu$ to preserve the holomorphic volume form on $X$.

If we work in a background where one of $\gamma^{1,0}$ or $\gamma^{1,2}$ is zero, then we see that $\mu^{1,1}$ is exactly a deformation of complex structure along $X$. 
In terms of the eleven-dimensional geometry, these field configurations describe deformations of the natural transverse holomorphic foliation (THF structure) on $X \times S$ which further preserve the holomorphic volume form along the leaves.
We further unpack the equations of motion in more general backgrounds in \S \ref{s:components}, but leave a complete study for future work.
%The equations of motion state that $\gamma$ is constant along $\RR$ and holomorphic along $\CC^5$. 
%Further, $\gamma$ is only defined up to the additional of a total holomorphic derivative $\gamma \sim \gamma + \del \beta$ where $\beta$ is a holomorphic function on $\CC^5$ which is constant along $\RR$.

@@

The simplest field configuration in the twisted theory is the flat background; this corresponds to considering to our eleven-dimensional theory on $\CC^5 \times \RR$, where we equip $\CC^5$ with the flat holomorphic volume form. 
In this case, we find a striking relationship to a certain infinite-dimensional simple super Lie algebra studied by Kac and collaborators \cite{KacBible,KacE510}.

\begin{thm}
The global symmetry algebra of the eleven-dimensional theory on $\CC^5 \times \RR$ is equivalent to a central extension of the exceptional simple super Lie algebra $E(5,10)$. 
\end{thm}

In particular, correlation functions involving observables of the eleven-dimensional theory will be constrained by the infinite-dimensional symmetry algebra $E(5,10)$. 
Given our main conjecture that the interacting eleven-dimensional BV theory on $\CC^5 \times \RR$ is the twist of supergravity, we obtain the following.

@@

Through the dimensional reduction of our eleven-dimensional theory we will find a match with Costello and Li's descriptions of twists of 10-dimensional supergravity in terms of BCOV theory.
In Table \ref{diagram}, we provide a summary of the comparisons with twists of supergravity in lower dimensions we perform. The squiggly arrows denote further twists, and the solid arrows denote various kinds of reductions.

\begin{table}
\begin{center}
\[
\begin{tikzcd}[row sep=large,column sep = 3ex]
                                                                         &  \parbox{3cm}{\centering minimal twist of 11d} \arrow[ld, "\text{slab }" '] \arrow[d, "\text{topological }S^1" description] \arrow[rdd, "\text{CY3}" description, pos=.7, bend left = 5] \arrow[rr, rightsquigarrow] \arrow[rd, "\text{holomorphic }S^1"] &                                                             &  \parbox{3cm}{\centering non-minimal twist of 11d} \arrow[d, "\text{topological }S^1" description] \\
 \parbox{3cm}{\centering minimal twist of type I}  &  \parbox{3cm}{\centering minimal twist of type IIA}\arrow[l, "\text{orbifold}", start anchor={[xshift=1.5ex]west}, end anchor={[xshift=-1.5ex]east}]\arrow[r, rightsquigarrow, crossing over, end anchor = {[xshift=2ex]west}]                                                                                                                                                                                                &  \parbox{3cm}{\centering SU(4) twist of type IIA} \arrow[r, rightsquigarrow, start anchor={[xshift=-2ex]east}, end anchor={[xshift=2ex]west}] &  \parbox{3cm}{\centering SU(2) twist of type IIA}                                                 \\
                                                                         &                                                                                                                                                                                                                                                          & \parbox{3cm}{\centering minimal twist of 5d $\cN=1$}                   &                                                                                                
\end{tikzcd}
\]
\end{center}
\caption{Relations between various twists}
\label{diagram}
\end{table}

In \S\ref{s:su4red} we will show that the reduction along $\{0\} \times \RR \times \{0\} \subset \CC^4 \times \CC \times \RR$ is equivalent to the $SU(4)$ twist of type IIA supergravity. 
The topological string approach does not lead to a description of the minimal, or holomorphic, twist of type IIA supergravity. 
In \S \ref{s:su5IIA}, we describe the reduction along the line $\{0\} \times \RR \subset \CC^5 \times \RR$ to obtain a conjectural description of the holomorphic $SU(5)$ twist of type IIA supergravity. 
In \S \ref{s:Ired}, we show that the reduction along the interval $\{0\}\times [0,1]\subset \CC^5\times [0,1]$ with certain boundary conditions on the endpoints of the interval is equivalent to the twist of type I supergravity. We further observe in \S \ref{s:orbifold} that the twist of type I arises as the fixed points of a natural $\ZZ/2$ action on the minimal twist of IIA. Next, in \S \ref{s:CY3} we study the reduction along a Calabi-Yau 3 fold to obtain a conjectural description of the minimal twist of 5d $\cN=1$ supergravity. Finally, we show in \S \ref{s:nonmin} how the non-minimal twist of eleven-dimensional supergravity arises as a further twist of our holomorphic twist.

%\subsection*{Future work}

\subsection*{Acknowledgements}
We warmly thank I.~Brunner, M.~Cederwall, K.~Costello, O.~Gwil\-liam, J.~Huerta, H.~Loosen, N.~Paquette, S.~R\"osch, J.~Walcher for conversation and inspiration of all kinds. I.S. and B.W. thank the University of Heidelberg, B.W. thanks the Mathematisches Forschungsinstitut Oberwolfach, and I.S. thanks the Krameterhof for hospitality while portions of this work were being completed. The work of S.R. is supported by the Perimeter Institute for Theoretical Physics. Research at Perimeter Institute is supported in part by the Government of Canada, through the Department of Innovation, Science and Economic Development Canada, and by the Province of Ontario, through the Ministry of Colleges and Universities.
The work of I.S. is supported by the Free State of Bavaria. The work of B.W. is supported by the University of Edinburgh.
% Any opinions, findings, and conclusions or recommendations expressed in this material are those of the authors and do not necessarily reflect the views of the funding agencies.
